From the confusion matrices below and the accuracy ratings we can see that both the random forest and the logistic linear regression produced similar accuracy ratings. The random forest once again had a significant dropoff from training data predicitions to test data, while the logistic model maintained its robustness. As for the Northwestern game, both the random forest and the logistic regression predicted a loss. It is safe to say you would not have made the same bet.


```{r}
library(tidyverse)
library(tidymodels)
library(zoo)
library(hoopR)

set.seed(1234)
```

```{r}
teamgames <- load_mbb_team_box(seasons = 2015:2022) %>%
  separate(field_goals_made_field_goals_attempted, into = c("field_goals_made","field_goals_attempted")) %>%
  separate(three_point_field_goals_made_three_point_field_goals_attempted, into = c("three_point_field_goals_made","three_point_field_goals_attempted")) %>%
  separate(free_throws_made_free_throws_attempted, into = c("free_throws_made","free_throws_attempted")) %>%
  mutate_at(12:34, as.numeric)

teamstats <- teamgames %>% 
  group_by(team_short_display_name, season) %>%
  arrange(game_date) %>%
  mutate(
    team_score = ((field_goals_made-three_point_field_goals_made) * 2) + (three_point_field_goals_made*3) + free_throws_made,
    true_shooting_percentage = (team_score / (2*(field_goals_attempted + (.44 * free_throws_attempted)))) * 100,
    turnover_pct = turnovers/(field_goals_attempted + 0.44 * free_throws_attempted + turnovers),
    free_throw_factor = free_throws_made/field_goals_attempted,
    team_rolling_true_shooting_percentage = rollmean(lag(true_shooting_percentage, n=1), k=10, align="right", fill=NA),
    team_rolling_turnover_percentage = rollmean(lag(turnover_pct, n=1), k=10, align="right", fill=NA),
    team_rolling_free_throw_factor = rollmean(lag(free_throw_factor, n=1), k=10, align="right", fill=NA),
    dummy_home_away = ifelse(home_away == "HOME", 1, 0)
  ) %>% ungroup()

opponent <- teamstats %>% select(game_id, team_id, offensive_rebounds, defensive_rebounds) %>% rename(opponent_id=team_id, opponent_offensive_rebounds = offensive_rebounds, opponent_defensive_rebounds=defensive_rebounds) %>% mutate(opponent_id = as.numeric(opponent_id))

newteamstats <- teamstats %>% 
  inner_join(opponent) %>% 
  mutate(
    orb = offensive_rebounds / (offensive_rebounds + opponent_defensive_rebounds),
    drb = defensive_rebounds / (opponent_offensive_rebounds + defensive_rebounds),
    team_rolling_orb = rollmean(lag(orb, n=1), k=10, align="right", fill=NA),
    team_rolling_drb = rollmean(lag(drb, n=1), k=10, align="right", fill=NA)
    )
```

```{r}
team_side <- newteamstats %>%
  select(game_id, team_id, team_short_display_name, opponent_id, game_date, season, team_score, team_rolling_true_shooting_percentage, team_rolling_free_throw_factor, team_rolling_turnover_percentage, team_rolling_orb, team_rolling_drb, dummy_home_away) %>% na.omit()

opponent_side <- newteamstats %>%
  select(game_id, team_id, team_short_display_name, team_score, team_rolling_true_shooting_percentage, team_rolling_free_throw_factor, team_rolling_turnover_percentage, team_rolling_orb, team_rolling_drb) %>% na.omit() %>%
  rename(
    opponent_id = team_id,
    opponent_short_display_name = team_short_display_name,
    opponent_score = team_score,
    opponent_rolling_true_shooting_percentage = team_rolling_true_shooting_percentage,
    opponent_rolling_free_throw_factor = team_rolling_free_throw_factor,
    opponent_rolling_turnover_percentage = team_rolling_turnover_percentage,
    opponent_rolling_orb = team_rolling_orb,
    opponent_rolling_drb = team_rolling_drb
  ) %>%
  mutate(opponent_id = as.numeric(opponent_id))

games <- team_side %>% inner_join(opponent_side) %>% mutate(
  TeamResult = as.factor(case_when(
    team_score > opponent_score ~ "W",
    opponent_score > team_score ~ "L"
))) %>% na.omit()
```

```{r}
games$TeamResult <- relevel(games$TeamResult, ref="W")

modelgames <- games %>% select(game_id, game_date, team_short_display_name, opponent_short_display_name, season, team_rolling_true_shooting_percentage, opponent_rolling_true_shooting_percentage, team_rolling_turnover_percentage, opponent_rolling_turnover_percentage, dummy_home_away, team_rolling_free_throw_factor, team_rolling_orb, team_rolling_drb, TeamResult) 
```



```{r}
game_split <- initial_split(modelgames, prop = .8)
game_train <- training(game_split)
game_test <- testing(game_split)
```

```{r}
game_recipe <- 
  recipe(TeamResult ~ ., data = game_train) %>% 
  update_role(game_id, game_date, team_short_display_name, opponent_short_display_name, season, new_role = "ID") %>%
  step_normalize(all_predictors())

summary(game_recipe)
```

```{r}
log_mod <- 
  logistic_reg() %>% 
  set_engine("glm") %>%
  set_mode("classification")

rf_mod <- 
  rand_forest() %>% 
  set_engine("ranger") %>%
  set_mode("classification")
```

```{r}
log_workflow <- 
  workflow() %>% 
  add_model(log_mod) %>% 
  add_recipe(game_recipe)

rf_workflow <- 
  workflow() %>% 
  add_model(rf_mod) %>% 
  add_recipe(game_recipe)
```

```{r}
log_fit <- 
  log_workflow %>% 
  fit(data = game_train)

rf_fit <- 
  rf_workflow %>% 
  fit(data = game_train)
```

```{r}
logpredict <- log_fit %>% predict(new_data = game_train) %>%
  bind_cols(game_train) 

logpredict <- log_fit %>% predict(new_data = game_train, type="prob") %>%
  bind_cols(logpredict)

rfpredict <- rf_fit %>% predict(new_data = game_train) %>%
  bind_cols(game_train) 

rfpredict <- rf_fit %>% predict(new_data = game_train, type="prob") %>%
  bind_cols(rfpredict)
```

```{r}
metrics(logpredict, TeamResult, .pred_class)
```

```{r}
metrics(rfpredict, TeamResult, .pred_class)
```

```{r}
logtestpredict <- log_fit %>% predict(new_data = game_test) %>%
  bind_cols(game_test)

logtestpredict <- log_fit %>% predict(new_data = game_test, type="prob") %>%
  bind_cols(logtestpredict)

metrics(logtestpredict, TeamResult, .pred_class)
```

```{r}
rftestpredict <- rf_fit %>% predict(new_data = game_test) %>%
  bind_cols(game_test)

rftestpredict <- rf_fit %>% predict(new_data = game_test, type="prob") %>%
  bind_cols(rftestpredict)

metrics(rftestpredict, TeamResult, .pred_class)
```


```{r}
logtestpredict %>%
  conf_mat(TeamResult, .pred_class)
```

```{r}
rftestpredict %>%
  conf_mat(TeamResult, .pred_class)
```

```{r}
nu <- modelgames %>% filter(team_short_display_name == "Nebraska" & season == 2022) %>% arrange(game_date)

nulogpredict <- log_fit %>% predict(new_data = nu) %>%
  bind_cols(nu)

nulogpredict <- log_fit %>% predict(new_data = nu, type="prob") %>%
  bind_cols(nulogpredict)

nulogpredict %>% select(.pred_class, TeamResult, opponent_short_display_name, everything())
```

```{r}
nurfpredict <- rf_fit %>% predict(new_data = nu) %>%
  bind_cols(nu)

nurfpredict <- rf_fit %>% predict(new_data = nu, type="prob") %>%
  bind_cols(nurfpredict)

nurfpredict %>% select(.pred_class, TeamResult, opponent_short_display_name, everything())
```

