#v5 updates:
  - implementing an exponential moving average of several important stats
  - particular implementation of 100 game ema's in key stats. Attempt to quantify team quality.
    - 100 game stats in some possibly key areas
  - removed the na_omit from everything before modelgames
  - omitted svm and xgboost to make simpler
  - added chunks after each operation to count NA's and see where they are coming from / maintain data quality
  
# metrics
  - 

```{r}
library(tidyverse)
library(tidymodels)
library(zoo)
library(hoopR)
library(corrr)
library(QuantTools)
```

```{r}
data <- load_mbb_team_box(seasons = 2022:2022) %>%
  separate(field_goals_made_field_goals_attempted, into = c("field_goals_made","field_goals_attempted")) %>%
  separate(three_point_field_goals_made_three_point_field_goals_attempted, into = c("three_point_field_goals_made","three_point_field_goals_attempted")) %>%
  separate(free_throws_made_free_throws_attempted, into = c("free_throws_made","free_throws_attempted")) %>%
  mutate_at(12:34, as.numeric) %>%
  rename(team = team_short_display_name,
         fgm = field_goals_made,
         threes = three_point_field_goals_made,
         ftm = free_throws_made,
         fta = free_throws_attempted,
         fga = field_goals_attempted,
         threes_attempted = three_point_field_goals_attempted,
         fg_pct = field_goal_pct
         ) %>% 
  arrange(game_date)

```

```{r}
# sums the nas in every column
na_count <- sapply(data, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
na <- data %>% filter(is.na(team))

# cleaning up data
data <- data %>% select(-team_alternate_color, -largest_lead, -team_logo, -team_color)
```


```{r}
  # cummean
  # look at spreads and see if you can beat them, maybe NBA project though

# important to group by in order to properly calculate rolling_means
teamdata <- data %>% 
  group_by(team, season) %>%
  # creating some new stats that we will use
  mutate(
    team_score = ((fgm - threes) * 2) + (threes*3) + ftm,
    possessions = fga - offensive_rebounds + turnovers + (.475 * fta),
    ppp = team_score/possessions,
    turnover_pct = turnovers/(fga + 0.44 * fta + turnovers),
    free_throw_factor = ftm/fga,
    true_shooting = (team_score / (2*(fga + (.44 * fta)))) * 100, 
    efg = (fgm + .5*threes)/fga,
    three_pct = threes/threes_attempted,
    three_pct = as.numeric(three_pct),
    # a binary variable to identify whether teams are home or away
    dummy_home_away = ifelse(home_away == "HOME", 1, 0),
    
    
    #rolling means, lag prevents model from knowing current results, align right is needed because rolling mean automatically chooses a value from the future without it, fill NA replaces errors with NA rather than the value of the game cell
    cum_three_pct = lag(cummean(three_pct), n = 1),
    cum_fg_pct = lag(cummean(fg_pct), n=1),
    cum_ppp = lag(cummean(ppp), n=1),
    cum_true_shoot = lag(cummean(true_shooting), n=1),
    cum_turnover_pct = lag(cummean(turnover_pct), n=1),
    cum_ftf = lag(cummean(free_throw_factor), n=1),
    cum_efg = lag(cummean(efg), n=1),
    
    r_three_pct = rollmean(lag(three_pct, n=1), k=10, align="right", fill=NA),
    r_fg_pct = rollmean(lag(fg_pct, n=1), k=10, align="right", fill=NA),
    r_ppp = rollmean(lag(ppp, n=1), k=10, align="right", fill=NA),
    r_true_shoot = rollmean(lag(true_shooting, n=1), k=10, align="right", fill=NA),
    r_turnover_pct = rollmean(lag(turnover_pct, n=1), k=10, align="right", fill=NA),
    r_ftf = rollmean(lag(free_throw_factor, n=1), k=10, align="right", fill=NA),
    r_efg = rollmean(lag(efg, n=1), k=10, align="right", fill=NA)
    ) %>% mutate(team_id = as.numeric(team_id))

excluded_vars <- c("team_uid", "team_display_name", "team_name", "team_location", "team_abbreviation", "team_slug", "opponent_mascot", "opponent_abbrev")
teamdata <- teamdata %>% select(-excluded_vars) %>% ungroup() 

# nu1 <- teamdata %>% filter(team == "Nebraska")
# view(nu1)
```

```{r}
# sums the nas in every column
na_count <- sapply(teamdata, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```


```{r}
opponentdata <- teamdata %>% #select(-excluded_vars) %>%
    rename(opponent = opponent_name,
          opp_id=opponent_id, 
          opp_offensive_rebounds = offensive_rebounds,
          opp_defensive_rebounds=defensive_rebounds,
    opp_score = team_score,
    opp_possessions = possessions, 
    opp_ppp = ppp, 
    opp_true_shooting = true_shooting, 
    opp_turnover_pct = turnover_pct,
    opp_fgm = fgm, 
    opp_fga = fga, 
    opp_fg_pct = fg_pct, 
    opp_threes = threes, 
    opp_threes_attempted = threes_attempted, 
    opp_three_pct = three_pct, 
    opp_ftm = ftm, 
    opp_fta = fta, 
    opp_free_throw_pct = free_throw_pct, 
    opp_free_throw_factor = free_throw_factor, 
    opp_total_rebounds = total_rebounds, 
    opp_offensive_rebounds = offensive_rebounds, 
    opp_defensive_rebounds = defensive_rebounds, 
    opp_assists = assists, 
    opp_steals = steals, 
    opp_blocks = blocks, 
    opp_turnovers = turnovers, 
    opp_fouls = fouls, 
    opp_three_point_field_goal_pct = three_point_field_goal_pct,
    opp_team_rebounds = team_rebounds,
    opp_team_turnovers = team_turnovers,
    opp_total_turnovers = total_turnovers,
    opp_total_technical_fouls = total_technical_fouls,
    opp_technical_fouls =technical_fouls,
    opp_flagrant_fouls = flagrant_fouls,
    opp_home_away = home_away,
    opp_efg = efg,
    opp_dummy_home_away = dummy_home_away,
    
    opp_cum_three_pct = cum_three_pct,
    opp_cum_true_shoot = cum_true_shoot,
    opp_cum_ppp = cum_ppp,
    opp_cum_fg_pct = cum_fg_pct,
    opp_cum_ftf = cum_ftf,
    opp_cum_turnover_pct = cum_turnover_pct,
    opp_cum_efg = cum_efg,
   
    opp_r_three_pct = r_three_pct,
    opp_r_true_shoot = r_true_shoot,
    opp_r_ppp = r_ppp,
    opp_r_fg_pct = r_fg_pct,
    opp_r_ftf = r_ftf,
    opp_r_turnover_pct = r_turnover_pct,
    opp_r_efg = r_efg,
    # opp_largestLead = largestLead
    )


excluded_vars <- c("opponent", "opp_id")
opponentdata <- opponentdata %>% select(-excluded_vars)
opponentdata <- opponentdata %>% rename(opponent_name = team,
                                        opponent_id = team_id)

# nu2 <- opponentdata %>% filter(opponent_id == 158)
# view(nu2)
```

```{r}
# sums the nas in every column
na_count <- sapply(opponentdata, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```


```{r}
bothdata <- teamdata %>% inner_join(opponentdata)

# nu3 <- bothdata %>% filter(team == "Nebraska")
# view(nu3)
```

```{r}
# sums the nas in every column
na_count <- sapply(bothdata, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```

```{r}
bothdata <- bothdata %>% group_by(team, season) %>% mutate(
    orb_pct = offensive_rebounds / (offensive_rebounds + opp_defensive_rebounds),
    drb_pct = defensive_rebounds / (opp_offensive_rebounds + defensive_rebounds),
    cum_orb_pct = lag(cummean(orb_pct), n=1),
    cum_drb_pct = lag(cummean(drb_pct), n=1),
    
    r_orb_pct = rollmean(lag(orb_pct, n=1), k=10, align="right", fill=NA),
    r_drb_pct = rollmean(lag(drb_pct, n=1), k=10, align="right", fill=NA),
    
    team_scoring_margin = team_score - opp_score,
    
    cum_score_margin = lag(cummean(team_scoring_margin), n=1),
    cum_def_ppg = lag(cummean(opp_score), n=1),
    cum_def_ppp = lag(cummean(opp_ppp), n=1),
    cum_def_three_pct = lag(cummean(opp_three_pct), n=1),
    cum_def_fg_pct = lag(cummean(opp_fg_pct), n=1),
    cum_def_true_shoot = lag(cummean(opp_true_shooting), n=1),
    cum_def_ftf = lag(cummean(opp_free_throw_factor), n=1),
    cum_def_efg = lag(cummean(opp_efg), n=1),
    cum_def_turnover_pct = lag(cummean(opp_turnover_pct), n=1), 
    
    r_score_margin = rollmean(lag(team_scoring_margin, n=1), k=10, align="right", fill=0),
    r_def_ppg = rollmean(lag(opp_score, n=1), k=10, align="right", fill=NA),
    r_def_ppp = rollmean(lag(opp_ppp, n=1), k=10, align="right", fill=NA),
    r_def_three_pct = rollmean(lag(opp_three_pct, n=1), k=10, align="right", fill=NA),
    r_def_fg_pct = rollmean(lag(opp_fg_pct, n=1), k=10, align="right", fill=NA),
    r_def_true_shoot = rollmean(lag(opp_true_shooting, n=1), k=10, align="right", fill=NA),
    r_def_ftf = rollmean(lag(opp_free_throw_factor, n=1), k=10, align="right", fill=NA),
    r_def_efg = rollmean(lag(opp_efg, n=1), k=10, align="right", fill=NA),
    r_def_turnover_pct = rollmean(lag(opp_turnover_pct, n=1), k=10, align="right", fill=NA),
    
    possession_diff = possessions - opp_possessions
    
    ) %>% ungroup()

# nu4 <- bothdata %>% filter(team == "Nebraska")
# view(nu4)
# osu4 <- bothdata %>% filter(team == "Ohio State")
# view(osu4)
```

```{r}
# sums the nas in every column
na_count <- sapply(bothdata, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```


```{r}
bothdata <- game100_emas <- bothdata %>% group_by(team) %>% mutate(win_loss = ifelse(team_score > opp_score, 1, 0),
                                                       game_100_win_pct_ema = lag(ema(win_loss, n=100), n=1),
                                                       game_100_ppp_ema = lag(ema(ppp, n=100), n=1),
                                                       game_100_orb_pct_ema = lag(ema(orb_pct, n=100), n=1),
                                                       game_100_drb_pct_ema = lag(ema(drb_pct, n=100), n=1),
                                                       game_100_opp_fg_pct_ema = lag(ema(opp_fg_pct, n=100), n=1),
                                                       game_100_turnover_pct_ema = lag(ema(turnover_pct, n=100), n=1),
                                                       game_100_ftf_ema = lag(ema(free_throw_factor, n=100), n=1),
                                                       game_100_true_shoot = lag(ema(true_shooting, n=100), n=1) 
                                                       ) %>% ungroup()
```

```{r}
# sums the nas in every column
na_count <- sapply(bothdata, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```


```{r}
opponent_rebound_data <- bothdata %>% select(game_id, team_id, game_date, season, orb_pct, drb_pct, r_orb_pct, r_drb_pct, team_scoring_margin, r_score_margin, r_def_ppg, r_def_ppp, r_def_three_pct, r_def_fg_pct, r_def_true_shoot, r_def_ftf, r_def_efg, r_def_turnover_pct, cum_orb_pct, cum_drb_pct, cum_score_margin, cum_def_ppg, cum_def_ppp, cum_def_three_pct, cum_def_fg_pct, cum_def_true_shoot, cum_def_ftf, cum_def_efg, cum_def_turnover_pct, possession_diff, game_100_win_pct_ema, game_100_ppp_ema, game_100_orb_pct_ema, game_100_drb_pct_ema, game_100_opp_fg_pct_ema, game_100_turnover_pct_ema, game_100_ftf_ema, game_100_true_shoot) %>%
  rename(opponent_id = team_id,
         opp_orb_pct = orb_pct,
         opp_drb_pct = drb_pct,
         
         opp_r_orb_pct = r_orb_pct,
         opp_r_drb_pct = r_drb_pct,
         opp_r_score_margin = r_score_margin,
    opp_r_def_ppg = r_def_ppg,
    opp_r_def_ppp = r_def_ppp,
    opp_r_def_three_pct = r_def_three_pct,
    opp_r_def_fg_pct = r_def_fg_pct,
    opp_r_def_true_shoot =r_def_true_shoot,
    opp_r_def_ftf = r_def_ftf,
    opp_r_def_efg = r_def_efg,
    opp_r_def_turnover_pct = r_def_turnover_pct,
    opp_possession_diff = possession_diff,
    
    opp_cum_orb_pct = cum_orb_pct,
    opp_cum_drb_pct = cum_drb_pct,
    opp_cum_score_margin = cum_score_margin,
    opp_cum_def_ppg = cum_def_ppg,
    opp_cum_def_ppp = cum_def_ppp,
    opp_cum_def_three_pct = cum_def_three_pct,
    opp_cum_def_fg_pct = cum_def_fg_pct,
    opp_cum_def_true_shoot = cum_def_true_shoot,
    opp_cum_def_ftf = cum_def_ftf,
    opp_cum_def_efg = cum_def_efg,
    opp_cum_def_turnover_pct = cum_def_turnover_pct,
    
    opp_game_100_win_pct_ema = game_100_win_pct_ema,
    opp_game_100_ppp_ema = game_100_ppp_ema,
    opp_game_100_orb_pct_ema = game_100_orb_pct_ema,
    opp_game_100_drb_pct_ema = game_100_drb_pct_ema,
    opp_game_100_opp_fg_pct_ema = game_100_opp_fg_pct_ema,
    opp_game_100_turnover_pct_ema = game_100_turnover_pct_ema,
    opp_game_100_ftf_ema = game_100_ftf_ema,
    opp_game_100_true_shoot = game_100_true_shoot, 
    
    
    opp_team_scoring_margin = team_scoring_margin
         )
bothdatawithrebounds <- bothdata %>% inner_join(opponent_rebound_data)

# nu5 <- bothdatawithrebounds %>% filter(team == "Nebraska")
# view(nu5)
```

```{r}
# sums the nas in every column
na_count <- sapply(opponent_rebound_data, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```

```{r}
# sums the nas in every column
na_count <- sapply(bothdatawithrebounds, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```



```{r}
teamdatawithreboundsanddefense <- bothdatawithrebounds %>% group_by(team, season) %>% mutate(score_margin = team_score - opp_score,
  score_vs_avg = score_margin + opp_r_score_margin,
  cum_score_vs_avg = lag(cummean(score_vs_avg), n=1),
  cum_possession_diff = lag(cummean(possession_diff), n=1),
  
  r_score_vs_avg = rollmean(lag(score_vs_avg, n=1), k=10, align="right", fill=NA),
  r_possession_diff = rollmean(lag(possession_diff, n=1), k=10, align="right", fill=NA),
) %>% ungroup()

# nu6 <- teamdatawithreboundsanddefense %>% filter(team == "Nebraska")
# view(nu6)
```

```{r}
# sums the nas in every column
na_count <- sapply(teamdatawithreboundsanddefense, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```

```{r}
opponent_defense_data <- teamdatawithreboundsanddefense %>% select(game_id, team_id, game_date, season, score_vs_avg, r_score_vs_avg, r_possession_diff, cum_score_vs_avg, cum_possession_diff) %>%
  rename(opponent_id = team_id,
         opp_score_vs_avg = score_vs_avg,
         opp_r_score_vs_avg = r_score_vs_avg,
         opp_r_possession_diff = r_possession_diff,
         
         opp_cum_score_vs_avg = cum_score_vs_avg,
         opp_cum_possession_diff = cum_possession_diff)
bothdatawithreboundsanddefense <- teamdatawithreboundsanddefense %>% inner_join(opponent_defense_data)

# nu7 <- bothdatawithreboundsanddefense %>% filter(team == "Nebraska")
# view(nu7)
# osu7 <- bothdatawithreboundsanddefense %>% filter(team == "Ohio State")
# view(osu7)
```

```{r}
# sums the nas in every column
na_count <- sapply(bothdatawithreboundsanddefense, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```


```{r}
games <- bothdatawithreboundsanddefense

games <- games %>% mutate(
  TeamResult = as.factor(case_when(
    team_score > opp_score ~ "W",
    opp_score > team_score ~ "L"
)))
```

```{r}
# sums the nas in every column
na_count <- sapply(games, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```

```{r}
modelgames <- games %>% select(game_id, game_date, team, opponent_name, season, dummy_home_away,
                               
                               r_true_shoot, opp_r_true_shoot, 
                               r_turnover_pct, opp_r_turnover_pct, 
                               r_ftf, opp_r_ftf, 
                               r_orb_pct, opp_r_orb_pct, 
                               r_fg_pct, opp_r_fg_pct, 
                               r_three_pct, opp_r_three_pct, 
                               r_ppp, opp_r_ppp,
                              
                               r_score_vs_avg, opp_r_score_vs_avg,
                               
                               r_def_ppg, opp_r_def_ppg,
                               r_def_ppp, opp_r_def_ppp,
                               r_def_three_pct, opp_r_def_three_pct,
                               r_def_fg_pct, opp_r_def_fg_pct,
                               r_def_true_shoot, opp_r_def_true_shoot,
                               r_def_ftf, opp_r_def_ftf,
                               r_def_efg, opp_r_def_efg,
                               r_def_turnover_pct, opp_r_def_turnover_pct,
                               r_possession_diff, opp_r_possession_diff,
                               
                               #cum stats
                               cum_true_shoot, opp_cum_true_shoot, 
                               cum_turnover_pct, opp_cum_turnover_pct, 
                               cum_ftf, opp_cum_ftf, 
                               cum_orb_pct, opp_cum_orb_pct, 
                               cum_fg_pct, opp_cum_fg_pct, 
                               cum_three_pct, opp_cum_three_pct, 
                               cum_ppp, opp_cum_ppp,
                               
                               cum_score_vs_avg, opp_cum_score_vs_avg,
                               
                               cum_def_ppg, opp_cum_def_ppg,
                               cum_def_ppp, opp_cum_def_ppp,
                               cum_def_three_pct, opp_cum_def_three_pct,
                               cum_def_fg_pct, opp_cum_def_fg_pct,
                               cum_def_true_shoot, opp_cum_def_true_shoot,
                               cum_def_ftf, opp_cum_def_ftf,
                               cum_def_efg, opp_cum_def_efg,
                               cum_def_turnover_pct, opp_cum_def_turnover_pct,
                               cum_possession_diff, opp_cum_possession_diff,
                               #r_blocks, opp_r_blocks,
                               
    opp_game_100_win_pct_ema, game_100_win_pct_ema,
    opp_game_100_ppp_ema, game_100_ppp_ema,
    opp_game_100_orb_pct_ema, game_100_orb_pct_ema,
    opp_game_100_drb_pct_ema, game_100_drb_pct_ema,
    opp_game_100_opp_fg_pct_ema, game_100_opp_fg_pct_ema,
    opp_game_100_turnover_pct_ema, game_100_turnover_pct_ema,
    opp_game_100_ftf_ema, game_100_ftf_ema,
    opp_game_100_true_shoot, game_100_true_shoot, 
                               
                               #rolling_sos, opponent_rolling_sos,
                               score_margin,
                               TeamResult) %>% na.omit()

```

```{r}
# sums the nas in every column
na_count <- sapply(modelgames, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
view(na_count)
# na <- data %>% filter(is.na(team))

# cleaning up data
# data <- data %>% select(-team_alternate_color, -largest_lead, team_logo, team_color)
```




```{r}
game_split <- initial_split(modelgames, prop = .8)
game_train <- training(game_split)
game_test <- testing(game_split)

game_recipe <- 
  recipe(TeamResult ~ ., data = game_train) %>% 
  update_role(game_id, game_date, team, opponent_name, season, score_margin, 
              new_role = "ID") %>%
  step_normalize(all_predictors())

summary(game_recipe)
```

```{r}
log_mod <- 
  logistic_reg() %>% 
  set_engine("glm") %>%
  set_mode("classification")
```

```{r}
rf_mod <- 
  rand_forest() %>% 
  set_engine("ranger") %>%
  set_mode("classification")
```

```{r}
log_workflow <- 
  workflow() %>% 
  add_model(log_mod) %>% 
  add_recipe(game_recipe)
```

```{r}
rf_workflow <- 
  workflow() %>% 
  add_model(rf_mod) %>% 
  add_recipe(game_recipe)
```

```{r}
log_fit <- 
  log_workflow %>% 
  fit(data = game_train)
```

```{r}
rf_fit <- 
  rf_workflow %>% 
  fit(data = game_train)
```

```{r}
logpredict <- log_fit %>% predict(new_data = game_train) %>%
  bind_cols(game_train) 

logpredict <- log_fit %>% predict(new_data = game_train, type="prob") %>%
  bind_cols(logpredict)
```

```{r}
rfpredict <- rf_fit %>% predict(new_data = game_train) %>%
  bind_cols(game_train) 

# rfpredict <- rf_fit %>% predict(new_data = game_train, type="prob") %>%
#   bind_cols(rfpredict)
```

```{r}
metrics(logpredict, TeamResult, .pred_class)
```

```{r}
metrics(rfpredict, TeamResult, estimate = .pred_class)
```

```{r}
logtestpredict <- log_fit %>% predict(new_data = game_test) %>%
  bind_cols(game_test)

logtestpredict <- log_fit %>% predict(new_data = game_test, type="prob") %>%
  bind_cols(logtestpredict)

metrics(logtestpredict, TeamResult, .pred_class)
```

```{r}
rftestpredict <- rf_fit %>% predict(new_data = game_test) %>%
  bind_cols(game_test)

# rftestpredict <- rf_fit %>% predict(new_data = game_test, type="prob") %>%
#   bind_cols(rftestpredict)

metrics(rftestpredict, TeamResult, estimate=.pred_class)
```